---
title: "Potential minima calculation"
output: html_document
---

The physical system is defined by the following potential energy:
$$
U=\sum_{k=1}^{N}r_k^2+\sum_{i,k=1;i>k}^{N}\frac{1}{r_{ik}^6} (1)
$$
Here N is the number of particles in the system, $r_k=\sqrt{x^2+y^2+z^2}$ is the radius-vector of a paticle number *k* in 3-dimensional space;  $r_{ik}=\sqrt{(x_i-x_k)^2+(y_i-y_k)^2+(z_i-z_k)^2}$ is the distance between particles *i* and *k* in 3-dimensional space.


We want to find equilibrium states of such system, which means that we are to find the minimum of the potential energy (1)  
We are going to do that by assigning N random radius-vectors for the initial state of the system, and then performing gradient descent calculations to define the potential minimums.  
First let's write a function to initialize the random state of the system:
```{r}
init<-function(N){##initializes matrix 3 x N with correct rownames and random values
        r<-cbind(rnorm(N),rnorm(N),rnorm(N))
        colnames(r)<-(c("x","y","z"))
        r
}
init(5)
```
We can see now the result of initialization for N=5.
In the future, we will store the state of the system in the matrices of same sort, for example, like this:
```{r}
r=init(5)
```
In the future we will need to be able to calculate the distance $r_{ki}$ between particles, so let's make a short function that does that:
```{r}
rki<-function(r,k,i){##calculates the distance between particle i and particle k; 
        ##x, y, z (vectors) specify coordinates of all particles
        rki<-NULL
        if(k==i){print("rki error! k=i!")}
        else{
                rki<-((r[k,1]-r[i,1])^2+(r[k,2]-r[i,2])^2+(r[k,3]-r[i,3])^2)^0.5
        }
        names(rki)<-"Distance"
        rki
}
```
Now lets check it out:
```{r}
distance=rki(r,2,3)
```
We can now see that the calculated distance is equal to `r distance`. We can check if its correct manually:
```{r}
r=matrix(c(1,2,1,2,1,2),ncol=3,nrow=2)
```










